---
title: 'Topic 7: Word Embeddings'
author: "Charles Hendrickson"
date: "05/17/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
```

This week's Rmd file here: <https://github.com/MaRo406/EDS_231-text-sentiment/blob/main/topic_7.Rmd>

```{r packages, include = FALSE}
library(here)
library(tidytext)
library(tidyverse)
library(widyr)
library(irlba) #singluar value decomposition
library(broom) # creating search_synonym function
library(textdata)
library(ggplot2)
library(dplyr)
library(data.table)

#https://semantle.com/
```

### Assignment

Download a set of pretrained vectors, GloVe, and explore them. 

Grab data here:
<!-- download.file('<https://nlp.stanford.edu/data/glove.6B.zip>',destfile = 'glove.6B.zip')  -->
<!-- unzip('glove.6B.zip')  -->
<!-- Use this file: 'glove.6B.300d.txt' -->

1.  Recreate the analyses in the last three chunks (find-synonyms, plot-synonyms, word-math) with the GloVe embeddings. How are they different from the embeddings created from the climbing accident data? Why do you think they are different?

2.  Run the classic word math equation, "king" - "man" = ?

3.  Think of three new word math equations. They can involve any words you'd like, whatever catches your interest.

```{r}
# read in the glove data
glove_data <- fread(here("../../data/glove.6B/glove.6B.300d.txt"), header = F)

# check if the data frame has row names 
has_rownames(glove_data)

# make a column into rownames
glove_data <- glove_data %>%
     column_to_rownames(var = 'V1')

# make a matrix 
glove_matrix <- data.matrix(glove_data)
```

```{r syn-function}
# create a function that searches for synonyms and produces similarity score
search_synonyms <- function(word_vectors, selected_vector) {
dat <- word_vectors %*% selected_vector
    
similarities <- dat %>%
        tibble(token = rownames(dat), similarity = dat[,1])

similarities %>%
       arrange(-similarity) %>%
        select(c(2,3))
}
```

```{r find-synonyms}
# use the search synonyms function to get the similarity scores for words like 'fall' and 'slip'.
fall <- search_synonyms(glove_matrix,glove_matrix["fall",])

slip <- search_synonyms(glove_matrix,glove_matrix["slip",])
```

**The similarity scores for the GloVe embeddings are much higher overall compared to the climbing accident embeddings. This difference could be due to the climbing data set containing text that exclusively uses the terms 'fall' and 'slip' to describe climbing situations instead of much generic terms such as 'decline' or 'wicket' meaning an opening like a window especially. These terms are very related to 'fall' and 'slip' when taken out of context, however the climbing data set does not use them. This difference could be due to the climbing data set being much smaller than the glove data set and also  climbing.**

```{r plot-synonyms}
slip %>%
    mutate(selected = "slip") %>%
    bind_rows(fall %>%
                  mutate(selected = "fall")) %>%
    group_by(selected) %>%
    top_n(15, similarity) %>%
    ungroup %>%
    mutate(token = reorder(token, similarity)) %>%
    ggplot(aes(token, similarity, fill = selected)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~selected, scales = "free") +
    coord_flip() +
    theme(strip.text=element_text(hjust=0, size=12)) +
    scale_y_continuous(expand = c(0,0)) +
    labs(x = NULL, title = "What word vectors are most similar to slip or fall?")
         
```

```{r word-math}
# word math equation, "king" - "man"
word_math1 <- glove_matrix["king",] - glove_matrix["man",] 
search_synonyms(glove_matrix, word_math1)

```

```{r}
# word math equation, "soldier" + "fighter"
word_math2 <- glove_matrix["soldier",] + glove_matrix["fighter",] 
search_synonyms(glove_matrix, word_math2)

```

```{r}
# word math equation, "flood" + "fill"
word_math3 <- glove_matrix["flood",] + glove_matrix["fill",] 
search_synonyms(glove_matrix, word_math3)

```

```{r}
# word math equation, "hunt" + "kill"
word_math4 <- glove_matrix["hunt",] + glove_matrix["kill",] 
search_synonyms(glove_matrix, word_math4)

```














